---
title: "Homework7.Rmd"
author: "Emiliano Saucedo, Pramit Vyas, Blake Underwood, Skyler Young, Zach Willens"
date: "2023-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(classdata)
data("happy", package = "classdata")
head(happy)
```

## question 1

```{r}
# replacing missing values with NA
HAPPY <- replace(happy, happy == "IAP", NA)
HAPPY <- replace(HAPPY, HAPPY == "DK", NA)
HAPPY <- replace(HAPPY, HAPPY == "NA", NA)
```

## question 2
```{r}
# change values to numeric
happy2 <- HAPPY %>% mutate(
  age = replace(age, age == "89 AND OLDER", 89),
  age = as.numeric(age)
)
```
## question 3
```{r}
levels(happy2$marital) = happy2$marital %>% fct_reorder(happy2$age, .fun = mean)
```


# Part 2

## Pramit Vyas
``` {r}
happy2 %>% drop_na(partyid) %>% drop_na(degree) %>% mutate(party = ifelse(grepl("dem", partyid), "Dem", ifelse(grepl("rep", partyid), "Rep", "Ind"))) %>% drop_na(sex, happy, health) %>% ggplot()+
    geom_mosaic(aes(x = product(degree, party), fill = degree), 
                na.rm = TRUE, divider =  mosaic("h"))+
    theme_bw()+
    theme(plot.title = element_text(hjust=0.5, size = rel(2)),
          axis.title.y = element_text(size = rel(1.8)),
          axis.title.x = element_text(size = rel(1.8)),
          axis.text.x = element_text(size = rel(1.8)),
          axis.text.y = element_text(size = rel(1.8)),
          strip.text.y = element_text(size = rel(1.8)))+
    labs(x = "Party", 
         y = "Education Level", 
         fill = "Education Level", 
         title = "Mosaic Plot of Party by Education Level") + facet_wrap(~happy)
```


It seems like in general the relationship between what party someone identifies as and hw much education they have remains constant throughout different happiness levels. Some key differences are that people that aren't too happy have lower education accross party lines, as fewer of them have gone to college. It also seems like republicans tend to by happier than democrats, meanwhile being independent doesn't seem to affect happiness. Also seems like for happier people, being more educated, such as getting a graduate degree means you are more likely to be democrat compared to the same contraints except less happy.

## Blake Underwood

```{r}
ggplot(happy) +
  aes(x = happy, fill = factor(degree)) +
  geom_bar(position = "fill")
```
We can clearly see from the porportions on the bar chart that those with higher level degrees such as Bachelors and Graduate tend to have higher porportions of "Very happy" people relative to their porportion of "Not happy" people. Those that didn't graduate high school have a higher proportion of "Not happy" relative to the porportion of "Very happy".

```{r}
ggplot(happy) +
  aes(x = happy, fill = factor(partyid)) +
  geom_bar(position = "fill")
```
The proportion of republicans increases from the "Not happy" category to the "Very happy" category. The Independent and democratic parties decrease in their porportions of "Not happy" to "Very happy". We can conclude that republicans self-report higher level of happiness then independents and democrats. 


## Zach Willens

```{r}
ggplot(happy) +
  aes(x = age, fill = factor(happy)) +
  geom_bar(position = "fill")
```
As a person ages we can see a slight increase in "very Happy", but as well to "Not too happy", which could signify that happiness is a devisive subject when one gets older in life. With the stereotypes of the "grumpy old person" coming to mind, as well as the "wise old man tales"


## Skyler Young

```{r}
ggplot(na.omit(HAPPY), aes(x = health, fill = factor(sex))) +
  geom_bar(position = "dodge", alpha = 0.6) +
  labs(title = "Distribution of perceived health by sex (excluding missing values)",
       x = "Perceived health",
       y = "Count",
       fill = "Sex") +
  scale_fill_discrete(name = "Sex", labels = c("Male", "Female"))
```
The plot displays the perceived health distribution by sex in the HAPPY dataset, excluding missing values. The x-axis depicts the levels of perceived health (1-5), while the y-axis shows the count of individuals in each group. Blue bars represent males, while orange bars represent females. More females than males are present in the dataset, and they tend to report slightly worse perceived health. Level 4 is the most common for both sexes, but females have a larger proportion at level 3. Overall, the plot suggests variation in the perceived health distribution by sex in the dataset.

##Emiliano Saucedo
```
ggplot(happy) +
  aes(x = happy, fill = factor(health)) +
  geom_bar(position = "fill")
```
Alot people get better health their happiness increases. According to the proportion the peoiple with mor health have the most happiness.
```
ggplot(happy) +
  aes(x = happy, fill = factor(finrela)) +
  geom_bar(position = "fill")
```
 
